---
title: "Diet Estimates for Realsies"
author: "Ashlee Mikkelsen"
date: "`r Sys.Date()`"
output: word_document
---

OK. I took the dry run with initial data exploration, TDF and prior calculations using Simmr. Now it's time to do the analysis for real. Use INLA to determine covariates, bring in MixSIAR and run the model for a real long time.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


First I load in my raw data and fix it up
```{r clean up and load packages, include=FALSE}

# clear workspace and load packages
rm(list = ls())
setwd("~/Rprojects/Diet-sensitivty-to-TDFs-and-priors")

library(readxl)
library(dplyr)
library(ggplot2)
library(viridis)
library(tidyverse)
library(MixSIAR)
library(INLA)
library(lattice)
library(nlme)  
library(plyr)
library(tidyr)
library(GGally)
library(mgcv)
library(brinla)

??inla.doc

```



```{r load data, include=FALSE}

BearDietSources <- read.csv("BearSources20220404.csv")
# remove 1 extreme outlier 
BearDietSources <- BearDietSources[-c(12),]

RawBearData_SI <- read_excel("RawBearData_SI_CORT.xlsx", 
    col_types = c("text", "text", "text", 
        "text", "text", "text", "text", "text", 
        "text", "text", "text", "text", "text", 
        "text", "text", "text", "date", "text", 
        "text", "text"))
str(RawBearData_SI)
warnings()

colSums(is.na(BearDietSources))

```



## Initial linear Analysis
### Data Cleanup

#### Fix bear isotope data
I need remove the records with missing values and correct the Carbon values for the Suess effect

```{r remove NAs}

bears.si <- subset(RawBearData_SI, N15!="NA")
bears.si <- subset(bears.si, year.represent!="NA")

bears.si$year.represent <- as.numeric(bears.si$year.represent)
bears.si$year.sampled <- as.numeric(bears.si$year.sampled)
bears.si$C13 <- as.numeric(bears.si$C13)
bears.si$N15 <- as.numeric(bears.si$N15)

LastYr <- max(bears.si$year.represent)

C13.suess <- bears.si$C13-(+0.022*(LastYr-bears.si$year.represent))
bears.si$C13.suess <- C13.suess

colSums(is.na(BearDietSources))

```

#### Fix Source Isotopes

Take raw SI data and convert moose hair to moose meat
```{r, include=FALSE}
##### First I will correct the moose end-members to represent moose meat and not moose hair
MooseH2M.N <- function(x,y){
  y= 0.88*x
}
MooseH2M.C <- function(x,y){
  y= 0.99*x
}
GenN <- function(x,y){
  y=1*x
}
GenC <- function(x,y){
  y=1*x
}
MooseMeatN <-rep(0,134)
MooseMeatC <- rep(0,134)
RawN <- BearDietSources$d15N
RawC <- BearDietSources$d13C
Dsource <- BearDietSources$Group
DF <- as.data.frame(cbind(Dsource,RawC,RawN,MooseMeatC,MooseMeatN))
DF$RawC <- as.numeric(DF$RawC)
DF$RawN <- as.numeric(DF$RawN)
DF$MooseMeatC <- as.numeric(DF$MooseMeatC)
DF$MooseMeatN <- as.numeric(DF$MooseMeatN)

for (i in 1:134) {
  if(DF$Dsource[i]=="Moose"){
    DF$MooseMeatN[i]=MooseH2M.N(DF$RawN[i])
  }else{
    DF$MooseMeatN[i]=GenN(DF$RawN[i])
  }
}

for (i in 1:134) {
  if(DF$Dsource[i]=="Moose"){
    DF$MooseMeatC[i]=MooseH2M.C(DF$RawC[i])
  }else{
    DF$MooseMeatC[i]=GenC(DF$RawC[i])
  }
}

BearDietSources$d13C.C <- DF$MooseMeatC
BearDietSources$d15N.C <- DF$MooseMeatN

``` 


### Data exploration

#### Sample Sizes
```{r}

table(bears.si$year.represent)
# 1992, 1993, and 2014 are problematic. I will remove 1992 & 1993, but leave 2014.
bears.si <- subset(bears.si, year.represent>1994)

table(bears.si$sex)
# Good balance of sexes, but I don't know what that NA is doing in there
bears.si <- subset(bears.si, sex!="NA")

table(bears.si$reprostatus.represent)
bears.si <- subset(bears.si, reprostatus.represent!="NA")


```


#### Outliers
```{r}

# Year is my only continuous variable
myvar <-as.matrix(cbind(bears.si[,5],bears.si[,13], bears.si[,21]))
dotchart(myvar)

dotchart(myvar[,1])
dotchart(myvar[,2])
dotchart(myvar[,3])

# No outliers to be concerned with

```


#### Relationships

```{r}

# carbon and Nitrogen
ggplot(data = bears.si, aes(C13.suess, N15))+
  geom_point()+
  geom_smooth()+
  theme_classic()

# year and Nitrogen
ggplot(data = bears.si, aes(year.represent, N15))+
  geom_point()+
  geom_smooth()+
  theme_classic()
# Well that is interesting... Do we see it in 13C?

#year and carbon
ggplot(data = bears.si, aes(year.represent, C13.suess))+
  geom_point()+
  geom_smooth()+
  theme_classic()
#different patter, but not a random distribution, a lot of annual variation in both of these. This will likely be an important random effect.


# sex and Nitrogen
ggplot(data = bears.si, aes(sex, N15))+
  geom_boxplot()+
  theme_classic()

#sex and carbon
ggplot(data = bears.si, aes(sex, C13.suess))+
  geom_boxplot()+
  theme_classic()

#reproduvtive status and Nitrogen
ggplot(data = bears.si, aes(reprostatus.represent, N15))+
  geom_boxplot()+
  theme_classic()

#reproduvtive status and carbon
ggplot(data = bears.si, aes(reprostatus.represent, C13.suess))+
  geom_boxplot()+
  theme_classic()
bears.si$fYR <- as.factor(bears.si$year.represent)
```



### INLA regression models

```{r}


C <- lm(C13.suess~1, data = bears.si)
summary(C)$sigma
priorpcC <- list(prec = list(prior = "pc.prec", param = c(0.35, 0.01)))

N <- lm(N15~1, data = bears.si)
summary(N)$sigma
priorpcN <- list(prec = list(prior = "pc.prec", param = c(0.70, 0.01)))

mCarbonrepro <- inla(C13.suess~reprostatus.represent+
                      f(BearID, model = "iid", hyper = priorpcC),
                    data = bears.si,
                    control.compute = list(dic = TRUE, waic = TRUE),
                    control.predictor = list(compute = TRUE),
                    quantiles = c(0.025, 0.975),
                    control.family = list(hyper = priorpcC)
                    )
mCarbonsex <- inla(C13.suess~sex+
                      f(BearID, model = "iid", hyper = priorpcC),
                    data = bears.si,
                    control.compute = list(dic = TRUE, waic = TRUE),
                    control.predictor = list(compute = TRUE),
                    quantiles = c(0.025, 0.975),
                    control.family = list(hyper = priorpcC)
                    )
summary(mCarbonrepro)
summary(mCarbonsex)

mNrepro <- inla(N15~reprostatus.represent+
                      f(BearID, model = "iid", hyper = priorpcN),
                    data = bears.si,
                    control.compute = list(dic = TRUE, waic = TRUE),
                    control.predictor = list(compute = TRUE),
                    quantiles = c(0.025, 0.975),
                    control.family = list(hyper = priorpcN)
                    )
mNsex <- inla(N15~sex+
                      f(BearID, model = "iid", hyper = priorpcN),
                    data = bears.si,
                    control.compute = list(dic = TRUE, waic = TRUE),
                    control.predictor = list(compute = TRUE),
                    quantiles = c(0.025, 0.975),
                    control.family = list(hyper = priorpcN)
                    )
summary(mNrepro)
summary(mNsex)

# So including repro status way out-performed sex. Therefore repro status should be included in the diet models

# Now to check year as continuous or factor

mCyear <- inla(C13.suess~reprostatus.represent+year.represent+
                      f(BearID, model = "iid", hyper = priorpcC),
                    data = bears.si,
                    control.compute = list(dic = TRUE, waic = TRUE),
                    control.predictor = list(compute = TRUE),
                    quantiles = c(0.025, 0.975),
                    control.family = list(hyper = priorpcC)
                    )
mCfYR <- inla(C13.suess~reprostatus.represent+fYR+
                      f(BearID, model = "iid", hyper = priorpcC),
                    data = bears.si,
                    control.compute = list(dic = TRUE, waic = TRUE),
                    control.predictor = list(compute = TRUE),
                    quantiles = c(0.025, 0.975),
                    control.family = list(hyper = priorpcC)
                    )
summary(mCyear)
summary(mCfYR)

mNyear <- inla(N15~reprostatus.represent+year.represent+
                      f(BearID, model = "iid", hyper = priorpcN),
                    data = bears.si,
                    control.compute = list(dic = TRUE, waic = TRUE),
                    control.predictor = list(compute = TRUE),
                    quantiles = c(0.025, 0.975),
                    control.family = list(hyper = priorpc)
                    )
mNfYR <- inla(N15~sex+fYR+
                      f(BearID, model = "iid", hyper = priorpcN),
                    data = bears.si,
                    control.compute = list(dic = TRUE, waic = TRUE),
                    control.predictor = list(compute = TRUE),
                    quantiles = c(0.025, 0.975),
                    control.family = list(hyper = priorpcN)
                    )
summary(mNyear)
summary(mNfYR)

# Ok, so using year as a factor way out-perfoms a continuous variable. So these two are the winners.

```



### Model Validation

#### Plot residuals versus fitted values
```{r}


head(mCfYR$summary.fitted.values)

F1 <- mCfYR$summary.fitted.values[,"mean"]
E1 <- bears.si$C13.suess-F1

df <- as.data.frame(cbind(F1,E1))

ggplot(aes(F1,E1), data = df)+
  geom_point()+
  geom_hline(yintercept = 0)+
  geom_smooth()+
  ylab("Residuals")+
  xlab("Posterior mean fitted 13C values")+
  theme_classic()
# This looks like we have a slight increase in variation in the residuals of the carbon values. The higher Carbon values would be bears eating the lingonberries..


head(mNfYR$summary.fitted.values)

F2 <- mNfYR$summary.fitted.values[,"mean"]
E2 <- bears.si$N15-F2

df2 <- as.data.frame(cbind(F2,E2))

ggplot(aes(F2,E2), data = df)+
  geom_point()+
  geom_hline(yintercept = 0)+
  geom_smooth()+
  ylab("Residuals")+
  xlab("Posterior mean fitted 15N values")+
  theme_classic()
# OOF. It looks like there is also something wacky and similar to the Carbon happening in the Nitrogen as well 


```


#### Plot residuals versus each covariate in the model

```{r}

# Carbon by repro status and fYR
boxplot(E1~fYR, data = bears.si)
abline(h=0)

boxplot(E1~reprostatus.represent, data = bears.si)
abline(h=0)


# Nitrogen by repro status and fYR
boxplot(E2~fYR, data = bears.si)
abline(h=0)

boxplot(E2~reprostatus.represent, data = bears.si)
abline(h=0)


```


#### Estimating effect size (importance)

```{r}


Beta1 <- mCfYR$summary.fixed[, c("mean",
                              "0.025quant",  
                              "0.975quant")] 
round(Beta1, digits = 3)

summary(mCarbonrepro)
summary(mCfYR)


Beta2 <- mNfYR$summary.fixed[, c("mean",
                              "0.025quant",  
                              "0.975quant")] 
round(Beta2, digits = 3)

summary(mNrepro)
summary(mNfYR)


```

#### Summary

There is a pretty strong year effect on both 13C and 15N. If MixSIAR can handle 2 random effects, year needs to be included. Otherwise I can nest reproductive status within year.


## Bayesian Mixing Model Anlaysis

#### Environment cleanup

I am now going to transition to estimating dietary proportions in MixSIAR using the covariates as determined above. To begin, I need to save my files and format them in a .csv identicle to the examples in the MixSIAR package to ensure they are useable to MixSIAR. Then I clear my R environment and load the data files.

```{r}

write.csv(bears.si,"BearIsotopes_Processed.csv")
write.csv(BearDietSources,"BearDieatSources_Cleaned.csv")

```

NOTE: In excel I removed all columns but those pertinent to the analysis: C13.Suess, N15, BearID, Reprostatus.represent, and year. represent. I renamed the isotope columns to match the MixSIAR standard. I also converted my covariates into categorical covariates with a dummy variable of numbers for each category. For example, repro.status recame repro where SolitaryF=1, SolitaryM=2,WithCubs=3, and WithMother=4. Year.represent was renamed yr with 1995=1 through 2015=20. BearID was changed to b1 through b378.






