---
title: "ComparingPriors"
author: "Ashlee Mikkelsen"
date: "`r Sys.Date()`"

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup
## WD and packages

```{r, include=FALSE}
##### Load in necessary data
setwd("~/Rprojects/Diet-sensitivty-to-TDFs-and-priors")

library(lme4)
library(MuMIn)
library(ggplot2)
library(viridis)
library(rstan)
library(brms)
library(wiqid)

```

## set graph theme

```{r, include=FALSE}
mytheme <- theme(
    axis.text = element_text(size = 11,face = "bold"),
    axis.title = element_text(size = 13, face = "bold"),
    panel.grid.major = element_line(color = "grey92"),
    panel.grid.minor = element_line(color = "grey96"),
    panel.background = element_rect(fill = "white"),
    axis.line = element_line(colour = "black",size = 1.2),
    axis.ticks = element_line(size = 1),
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 11, face = "bold")
    )
theme_set(mytheme)
```

# Data

```{r}
Mixture.RAW <- read.csv("RawBearData_SI_CORT.csv")
BearDietSources <- read.csv("BearDietSources_Cleaned.csv")

```


##### Take raw SI data and convert moose hair to moose meat and correct for Suess
```{r, include=FALSE}
##### First I will correct the moose end-members to represent moose meat and not moose hair
MooseH2M.N <- function(x,y){
  y= 0.88*x
}
MooseH2M.C <- function(x,y){
  y= 0.99*x
}
GenN <- function(x,y){
  y=1*x
}
GenC <- function(x,y){
  y=1*x
}
MooseMeatN <-rep(0,133)
MooseMeatC <- rep(0,133)
RawN <- BearDietSources$d15N
RawC <- BearDietSources$d13C
Dsource <- BearDietSources$Group
DF <- as.data.frame(cbind(Dsource,RawC,RawN,MooseMeatC,MooseMeatN))
DF$RawC <- as.numeric(DF$RawC)
DF$RawN <- as.numeric(DF$RawN)
DF$MooseMeatC <- as.numeric(DF$MooseMeatC)
DF$MooseMeatN <- as.numeric(DF$MooseMeatN)

for (i in 1:133) {
  if(DF$Dsource[i]=="Moose"){
    DF$MooseMeatN[i]=MooseH2M.N(DF$RawN[i])
  }else{
    DF$MooseMeatN[i]=GenN(DF$RawN[i])
  }
}

for (i in 1:133) {
  if(DF$Dsource[i]=="Moose"){
    DF$MooseMeatC[i]=MooseH2M.C(DF$RawC[i])
  }else{
    DF$MooseMeatC[i]=GenC(DF$RawC[i])
  }
}

BearDietSources$d13C.C <- DF$MooseMeatC
BearDietSources$d15N.C <- DF$MooseMeatN

```


## Clean data
```{r, include=FALSE}
bear.si <- subset(Mixture.RAW, N15!="NA")
bear.si <- subset(bear.si, year.represent!="NA")
LastYr <- as.numeric(max(bear.si$year.represent))

bear.si$year.represent <- as.numeric(bear.si$year.represent)
bear.si$C13 <- as.numeric(bear.si$C13)

C13.suess <- bear.si$C13-(+0.022*(LastYr-bear.si$year.represent))
bear.si$C13.suess <- C13.suess
```


# Side bar: N by sex and size

```{r, include=FALSE}
DF <- subset(bear.si,reprostatus.represent!="WithMother")
DF <- na.omit(DF)
colSums(is.na(DF))
DF$HCsq <- DF$headcirc^2
DF$HCln <- log(DF$headcirc)
DF$fYR <- as.factor(DF$year.represent)
DF$N15ln <- log(DF$N15)

DF$ZHC <- standardize(DF$headcirc)
DF$ZHCsq <- standardize(DF$HCsq)
DF$ZHCln <- standardize(DF$HCln)
colSums(is.na(DF$N15))

ggplot(data = DF, aes(N15))+
  geom_histogram(binwidth = 0.5*sd(DF$N15))

ggplot(data = DF, aes(N15ln))+
  geom_histogram(binwidth = 0.5*sd(DF$N15ln))


Sizebrm1 <- brm(N15 ~1+(1|BearID),
               data = DF,
               family = gaussian(link = "log"),
              iter = 8000,
              thin = 10)
summary(Sizebrm1)
options(mc.cores = 1)
fit1 <- add_criterion(Sizebrm1, criterion = "loo")
gc()
print(fit1$criteria$loo)
Int.LOOic <- fit1$criteria$loo$estimates[3,1]


Sizebrm2 <- brm(N15 ~ZHC+(1|BearID),
               data = DF,
               family = gaussian(link = "log"),
              iter = 8000,
              thin = 10)
summary(Sizebrm2)
options(mc.cores = 1)
fit2 <- add_criterion(Sizebrm2, criterion = "loo")
gc()
print(fit2$criteria$loo)
HC.LOOic <- fit2$criteria$loo$estimates[3,1]


Sizebrm3 <- brm(N15 ~ZHC+ZHCsq+(1|BearID),
               data = DF,
               family = gaussian(link = "log"),
              iter = 8000,
              thin = 10)
summary(Sizebrm3)
options(mc.cores = 1)
fit3 <- add_criterion(Sizebrm3, criterion = "loo")
gc()
print(fit3$criteria$loo)
HCsq.LOOic <- fit3$criteria$loo$estimates[3,1]

Sizebrm4 <- brm(N15 ~ZHCln+(1|BearID),
               data = DF,
               family = gaussian(link = "log"),
              iter = 8000,
              thin = 10)
summary(Sizebrm4)
options(mc.cores = 1)
fit4 <- add_criterion(Sizebrm4, criterion = "loo")
gc()
print(fit4$criteria$loo)
HCln.LOOic <- fit4$criteria$loo$estimates[3,1]


Sizebrm5 <- brm(N15 ~sex+ZHCln+(1|BearID),
               data = DF,
               family = gaussian(link = "log"),
              iter = 8000,
              thin = 10)
summary(Sizebrm5)
options(mc.cores = 1)
fit5 <- add_criterion(Sizebrm5, criterion = "loo")
gc()
print(fit5$criteria$loo)
SexHCln.LOOic <- fit5$criteria$loo$estimates[3,1]

Sizebrm6 <- brm(N15 ~sex+ZHCln+sex:ZHCln+(1|BearID),
               data = DF,
               family = gaussian(link = "log"),
              iter = 8000,
              thin = 10)
summary(Sizebrm6)
options(mc.cores = 1)
fit6 <- add_criterion(Sizebrm6, criterion = "loo")
gc()
print(fit5$criteria$loo)
SexHClninx.LOOic <- fit6$criteria$loo$estimates[3,1]
```


```{r, include=FALSE}
##### Females only, as Anne thought the trend was not supported in Fems
FemsOnly <- subset(DF, sex!="M")
SizeFems1lm <- lmer(N15~1+(1|BearID), data = FemsOnly)
SizeFems2lm <- lmer(N15~headcirc+(1|BearID), data = FemsOnly)

FemsAIC1 <- AICc(SizeFems1lm)
FemsAIC2 <- AICc(SizeFems2lm)
FemsAIC2-FemsAIC1

summary(SizeFemslm)

########
Males <- subset(DF, sex=="M")
Females <- subset(DF, sex=="F")

M15N <- function(x){
  exp(1.63+(0.07*x))
}

Zmhc <- seq(min(Males$ZHCln),max(Males$ZHCln),
           length.out=200)

Mnitro <- M15N(Zmhc)
Mnitro
aMHC <- exp((Zmhc*sd(DF$HCln))+mean(DF$HCln))
aMHC

F15N <- function(x){
  exp(1.51+(0.02*x))
}
Zfhc <- seq(min(Females$ZHCln),max(Females$ZHCln),
           length.out=200)
aFHC <- exp((Zfhc*sd(DF$HCln))+mean(DF$HCln))

Fnitro <- F15N(Zfhc)
Fnitro

Nitro.HC <- as.data.frame(cbind(aMHC,Mnitro,aFHC,Fnitro))

P <- palette(viridis(12))
P <- palette(viridis(12))


pSB1 <- ggplot(data = Nitro.HC)+
  geom_jitter(data = DF, aes(headcirc, N15, color=sex),
              width = 0.2, size=2, alpha=0.35)+
  geom_line(aes(aMHC,Mnitro), lwd=1.2, color=P[3])+
  geom_line(aes(aFHC,Fnitro), lwd=1.2, color=P[9])+
  scale_color_manual(values = c(P[3], P[9]),
                     breaks =c("M","F"))+
  xlab("Skull circumference (cm)")+
  ylab(expression(paste(delta^15, "N (\u2030)", sep = "")))+
  scale_y_continuous(expand = c(0,0))+
  scale_x_continuous(expand = c(0,0))+
  labs(color="Sex")
  
tiff("SexHeadCircLM.tiff", units="in", width=7, height=5, res=800)
pSB1
dev.off()

Agelm1 <- lmer(N15~age+sex+(1|BearID), data = DF)
AgeAICc1 <- AICc(Agelm1)
summary(Agelm1)

Agelm2 <- lmer(N15~age+sex+age:sex+(1|BearID), data = DF)
AgeAICc2 <- AICc(Agelm2)
summary(Agelm2)

```



```{r load and prep data load rquired packages, include=FALSE}

library(simmr)
library(readxl)
library(dplyr)
library(ggplot2)
library(viridis)

##### Taking data and turning it into a form usable by  the Simmr package
bearmix <- cbind(bear.si$C13.suess,bear.si$N15)
colnames(bearmix) = c('d13C','d15N')

bearsourcres_species <-BearDietSources %>%
  group_by(Group)%>%
  summarise(
    mean13C= mean(d13C),
    mean15N=mean(d15N),
    sd13C=sd(d13C),
    sd15N=sd(d15N),
    mean13Cconc=mean(PrctC),
    mean15Nconc=mean(PrctN)
  )
bears_species_names <- matrix(c("Ants","Bilberry","Crowberry","Lingonberry","Moose"))
bears_species_means <- bearsourcres_species[,2:3]
bears_species_sds <- as.matrix(bearsourcres_species[,4:5])
bear_species_conc <- as.matrix((bearsourcres_species[,6:7])/100)
bear_speciesc_sds <- rep(1,10)
bear_speciesc_sds <- matrix(bear_speciesc_sds, nrow = 5,ncol = 2)


```



```{r calculate TDFs for diet estimation, include=FALSE}

Mikkelsen13C <- function(x,y){
  y=-10.6+(0.42*x)
}

Mikkelsen15N <- function(x,y){
  y=5.02+(0.90*x)
}

# I use these linear equations to predict the isotopic signature of a brown bear eating 100% of a diet source (y) given the mean isotopic signature of that source (x)


d13C.Mikkelsen <- Mikkelsen13C(bears_species_means$mean13C)
print(d13C.Mikkelsen)

# The actual TDF is the difference between our predicted isotopic signature for a bear eating 100% of that source and the mean isotopic signature for that source
C13TDF.Mikkelsen <-abs(d13C.Mikkelsen-bears_species_means$mean13C)

# Here are the TDFs for Carbon for ants, bilberries, crowberries, lingonberries, and moose
print(C13TDF.Mikkelsen)

# Now we repeat the process for Nitrogen

d15N.Mikkelsen <- Mikkelsen15N(bears_species_means$mean15N)
print(d15N.Mikkelsen)
N15TDF.Mikkelsen <-d15N.Mikkelsen-bears_species_means$mean15N

# And here we have the TDFs for Nitrogen for ants, bilberries, crowberry, lingonberry, and moose
print(N15TDF.Mikkelsen)

bear_speciesTDF_means.Mikkelsen <- cbind(C13TDF.Mikkelsen,N15TDF.Mikkelsen)
print(bear_speciesTDF_means.Mikkelsen)

bears_species_means <- as.matrix(bearsourcres_species[,2:3])


```

# Introduction
The final step in understanding bear dietary estimates is looking at the effect of priors. There is disagreement among ecologists that use Bayesian frameworks about the use of informitive priors. I looked at three models, one with uninformitive priors, one with informitive priors, and one with very informitive priors. I then compared the distribution of the prior distribution to the posterior distribution and changes in the resulting diet estimates from each of the models.

## Sample summary

```{r}




```



```{r bear diet estimates with uninformitive priors, include=FALSE}

library(simmr)

bearsimmr_TDF_Mikkelsen = simmr_load(
  mixtures=bearmix,
  source_names=bears_species_names,
  source_means=bears_species_means,
  source_sds=bears_species_sds,
  correction_means=bear_speciesTDF_means.Mikkelsen,
  correction_sds=bear_speciesc_sds,
  concentration_means = bear_species_conc
  )

p1 <- plot(bearsimmr_TDF_Mikkelsen,
           xlab = expression(paste(delta^13, "C (\u2030)", sep = "")),
           ylab = expression(paste(delta^15, "N (\u2030)", sep = "")),
           mix_name = "Brown bears",
           title = ""
           )
p1 <- p1+
  scale_shape_manual(values = c(1,15,16,17,18,4))+
  theme(axis.text = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 23, face = "bold"),
        legend.text = element_text(size = 18))+
  scale_y_continuous(breaks = seq(-4,9,1),
                     labels = c("","","-2","","0","","2","",
                                "4","","6","","8",""))+
  coord_equal()

p1

tiff("IsoSpace.tiff", units="in", width=8, height=12, res=600)
p1
dev.off()

```

## Model with uninformitive priors
To compare the effect of adding informitive priors, I first ran a model with uninformitive priors. For this model I used the trophic discrimination factors that I derived using Hilerbrand et al. (1996), Felicetti et al. (2003), and Rode et al. (2016). The input for this model is the same used in determining groups and in comparing different trophic discrimination factors.

```{r run Simmr uninformitive priors, include=FALSE}

bearsimmr_UnPrior_Mikkelsen.out = simmr_mcmc(
  bearsimmr_TDF_Mikkelsen,
  mcmc_control = list(
    iter=1000000, burn=500000, thin=500, n.chain=4
  )
)


```



```{r Uninformitive priors model diagnostics and summary, echo=FALSE}
library(ggplot2)
library(viridis)
summary(bearsimmr_UnPrior_Mikkelsen.out, type = "diagnostics")

summary(bearsimmr_UnPrior_Mikkelsen.out, type="statistics")
summary(bearsimmr_UnPrior_Mikkelsen.out,'quantiles')
plot(bearsimmr_UnPrior_Mikkelsen.out, type = "density")

plot(bearsimmr_UnPrior_Mikkelsen.out, type = "matrix")

p2 <- plot(bearsimmr_UnPrior_Mikkelsen.out,
  type = "boxplot",
  title = ""
)+
  theme(
    axis.text = element_text(size = 18, face = "bold"),
    axis.title = element_text(size = 20, face = "bold")
  )+
  xlab("Diet Source")

p2

tiff("EstBoxPlot.tiff", units="in", width=15, height=7, res=800)
p2
dev.off()


DietEst.NIpriors <- as.data.frame(bearsimmr_UnPrior_Mikkelsen.out$output$`1`$BUGSoutput$sims.list$p)

write.csv(DietEst.NIpriors, "DietEst.NIpriors.csv")

DietEst.NIpriors <- read.csv("DietEst.NIpriors.csv")

pal <- palette(viridis(5))


p3 <- ggplot(data = DietEst.NIpriors)+
  geom_density(aes(Ants), color= pal[1], fill= pal[1], alpha=0.6, lwd=1)+
  geom_density(aes(Bilberry), color= pal[2], fill= pal[2], alpha=0.6, lwd=1)+
  geom_density(aes(Crowberry), color= pal[3], fill= pal[3], alpha=0.6, lwd=1)+
  geom_density(aes(Lingonberry), color= pal[4], fill= pal[4], alpha=0.6, lwd=1)+
  geom_density(aes(Moose), color= pal[5], fill= pal[5], alpha=0.6, lwd=1)+
  xlab("Diet Proportions")+
  ylab("Density")+
  theme(
    axis.text = element_text(size = 18,face = "bold"),
    axis.title = element_text(size = 20, face = "bold"),
    panel.grid.major = element_line(color = "grey92"),
    panel.grid.minor = element_line(color = "grey96"),
    panel.background = element_rect(fill = "white"),
    axis.line = element_line(colour = "black",size = 1),
    axis.ticks = element_line(size = 1),
    )+
  scale_y_continuous(expand = c(0,0), breaks = seq(0,200,25))+
  geom_text(label="Moose", x=0.06, y=135, size=8)+
  geom_text(label="Ants", x=0.085, y=80, size=8)+
  geom_text(label="Crowberry", x=0.157, y=20, size=8)+
  geom_text(label="Lingonberry", x=0.3, y=20, size=8)+
  geom_text(label="Bilberry", x=0.55, y=30, size=8)
p3

tiff("EstDenPlot.tiff", units="in", width=15, height=7, res=800)
p3
dev.off()


```

This graph compares the density of the prior distribution to the posterior distribution

```{r compare uninformitive priors to estimates, echo=FALSE}

prior_viz(bearsimmr_UnPrior_Mikkelsen.out)

```
## informitive priors

Now I calculate informitive priors. These priors are the mean dietary proportions based on estimated dietary energy content from fecal analysis of brown bears in Sweden and Norway. These studies calculated the proportions of more diet sources than in this analysis so ants, bilberries, crowberry, lingonberry, and moose only make up 0.89 of Scandinavian bears. Because the proportions must sum to 1, the remaining proportion of the diet was added into my five diet sources, distributed evenly to maintain ratios. The standard deviations were all around 0.10, except for moose, which had a standard deviation of 0.16, which I incorporated into the variance around my prior estimates.

```{r calculate priors with Simmr, echo=FALSE}

InformitivePriors <- read_excel("PriorCalculations.xlsx")

bearprior=simmr_elicit(n_sources = 5,
                   c(0.20,0.19,0.16,0.09,0.36),
                   c(0.04,0.04,0.04,0.04,0.02))

```


```{r run model with informitive priors, include=FALSE}

bearsimmr_Priors.out <- simmr_mcmc(
  bearsimmr_TDF_Mikkelsen,
  prior_control = list(
    means=bearprior$mean,
    sd=bearprior$sd
  ),
  mcmc_control = list(
    iter=1000000, burn=500000, thin=500, n.chain=4
  )
)


```

We can compare the estimates and variance around the estimates that we get when we use informative priors versus uninformative priors.

```{r diagnostics and summary of informitive priors, echo=FALSE}

summary(bearsimmr_Priors.out, type = "diagnostics")

summary(bearsimmr_Priors.out, type="statistics")
summary(bearsimmr_Priors.out,'quantiles')
plot(bearsimmr_Priors.out, type = "density")

plot(bearsimmr_Priors.out, type = "matrix")


p4 <- plot(bearsimmr_Priors.out,
  type = "boxplot",
  title = ""
)+
  theme(
    axis.text = element_text(size = 18, face = "bold"),
    axis.title = element_text(size = 20, face = "bold")
  )+
  xlab("Diet Source")

p4

tiff("EstBoxPlot_InPrior.tiff", units="in", width=15, height=7, res=800)
p4
dev.off()


DietEst.INpriors <- as.data.frame(bearsimmr_Priors.out$output$`1`$BUGSoutput$sims.list$p)
DietEst.INpriors$sim <- seq(1,4000)

ggplot(data = DietEst.INpriors, aes(sim,Moose))+
  geom_point()

ggplot(data = DietEst.INpriors, aes(Moose))+
  geom_histogram(bins = 30)+
   stat_bin(aes(y=..count.., label=..count..), geom="text", vjust=-.5)
(1+6+7+10+21+18+29+19+20+14+10+3)/(621+8+1+1)


pal <- palette(viridis(5))

mytheme <- theme(
    axis.text = element_text(size = 18,face = "bold"),
    axis.title = element_text(size = 20, face = "bold"),
    panel.grid.major = element_line(color = "grey92"),
    panel.grid.minor = element_line(color = "grey96"),
    panel.background = element_rect(fill = "white"),
    axis.line = element_line(colour = "black",size = 1),
    axis.ticks = element_line(size = 1),
    )

p5 <- ggplot(data = DietEst.INpriors)+
  geom_density(aes(Ants), color= pal[1], fill= pal[1], alpha=0.7)+
  geom_density(aes(Bilberry), color= pal[2], fill= pal[2], alpha=0.7)+
  geom_density(aes(Crowberry), color= pal[3], fill= pal[3], alpha=0.7)+
  geom_density(aes(Lingonberry), color= pal[4], fill= pal[4], alpha=0.7)+
  geom_density(aes(Moose), color= pal[5], fill= pal[5], alpha=0.7)+
  xlab("Diet Proportions")+
  ylab("Density")+
  mytheme+
  scale_y_continuous(expand = c(0,0), breaks = seq(0,200,25))+
  geom_text(label="Moose", x=0.095, y=100, size=8)+
  geom_text(label="Ants", x=0.022, y=70, size=8)+
  geom_text(label="Crowberry", x=0.145, y=18, size=8)+
  geom_text(label="Lingonberry", x=0.38, y=15, size=8)+
  geom_text(label="Bilberry", x=0.555, y=22, size=8)+
  labs(fill="Diet Source")

p5


tiff("EstDenPlot_InPriors.tiff", units="in", width=15, height=7, res=800)
p5
dev.off()



```

This gives the prior and posterior distributions from each model


```{r comapring estimates and distributions of uninformitive vs informitive priors}


p6 <- prior_viz(bearsimmr_Priors.out)


tiff("EstDenPriPostPlot_InPriors.tiff", units="in", width=15, height=7, res=800)
p6
dev.off()



```

The distribution of my priors are fairly wide and is only similar to the posterior distribution in crowberry. It looks like these priors have a small effect on my diet estimates, and that the prior distributions based on scat analysis is quite different from the estimates derived from stable isotopes.

```{r less informitive priors}


bearprior2=simmr_elicit(n_sources = 5,
                   c(0.20,0.19,0.16,0.09,0.36),
                   c(0.1,0.1,0.1,0.1,0.1))
```

## Less-informitive priors

Because the addition of my informitive priors had only small effects on my diet estimates, I included another set of priors with standard deviations half of large as I used for the some-what informitive priors.

```{r}

bearsimmr_LessInformPriors.out <- simmr_mcmc(
  bearsimmr_TDF_Mikkelsen,
  prior_control = list(
    means=bearprior2$mean,
    sd=bearprior2$sd
  ),
  mcmc_control = list(
    iter=1000000, burn=500000, thin=500, n.chain=4
  )
)


```


```{r}

summary(bearsimmr_LessInformPriors.out, type = "diagnostics")

summary(bearsimmr_LessInformPriors.out, type="statistics")
summary(bearsimmr_LessInformPriors.out,'quantiles')
plot(bearsimmr_LessInformPriors.out, type = "density")
plot(bearsimmr_LessInformPriors.out, type = "matrix")


```


```{r}

prior_viz(bearsimmr_TDF_Mikkelsen.out)
prior_viz(bearsimmr_Priors.out)
prior_viz(bearsimmr_LessInformPriors.out)

save.image("DietModelOutputs")

```




```{r}

Fecal.SI <- read_excel("Fecal_StableIsotope_estimates.xlsx")

ggplot(data = Fecal.SI, aes(Model,Estimate))+
  geom_col(aes(group=DietSource, fill=DietSource))+
  scale_fill_viridis(discrete = TRUE)+
  theme(axis.text = element_text(size=20),
        axis.title = element_text(size=20,face = "bold"),
        panel.background = element_rect(fill = "grey97"),
        legend.text = element_text(size=20),
        legend.title = element_text(size=20,face="bold"))+
  ylab("Estimated diet proportion")
  

```

# Look at differences in diets between males and females

### First, females


```{r}

bear.fem <- subset(bear.si,sex!="M")
bear.fem <- subset(bear.fem, reprostatus.represent!="WithMother")

ggplot(data = bear.fem, aes(N15, headcirc))+
  geom_jitter()+
  mytheme+
  geom_smooth(method = "lm")

library(lme4)
Fem.lm <- lmer(headcirc ~N15+(1|BearID), data = bear.fem)
summary(Fem.lm)

1.45-(1.96*0.59)


Fbearmix <- cbind(bear.fem$C13.suess,bear.fem$N15)
colnames(Fbearmix) = c('d13C','d15N')

bearsimmr_Fems = simmr_load(
  mixtures=Fbearmix,
  source_names=bears_species_names,
  source_means=bears_species_means,
  source_sds=bears_species_sds,
  correction_means=bear_speciesTDF_means.Mikkelsen,
  correction_sds=bear_speciesc_sds,
  concentration_means = bear_species_conc
  )

p10 <- plot(bearsimmr_Fems,
           xlab = expression(paste(delta^13, "C (\u2030)", sep = "")),
           ylab = expression(paste(delta^15, "N (\u2030)", sep = "")),
           mix_name = "Brown bears",
           title = ""
           )+
  scale_shape_manual(values = c(1,15,16,17,18,4))+
  theme(axis.text = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 23, face = "bold"),
        legend.text = element_text(size = 18),
        legend.title = element_blank())+
  scale_y_continuous(breaks = seq(-4,9,1),
                     labels = c("","","-2","","0","","2","",
                                "4","","6","","8",""))+
  coord_equal()

p10

bearsimmr_UnPrior_Fems = simmr_mcmc(
  bearsimmr_Fems,
  mcmc_control = list(
    iter=1000000, burn=500000, thin=500, n.chain=4
  )
)

summary(bearsimmr_UnPrior_Fems, type = "diagnostics")

summary(bearsimmr_UnPrior_Fems, type="statistics")
summary(bearsimmr_UnPrior_Fems,'quantiles')
plot(bearsimmr_UnPrior_Fems, type = "density")
plot(bearsimmr_UnPrior_Fems, type = "matrix")

p12 <- plot(bearsimmr_UnPrior_Fems,
  type = "boxplot",
  title = ""
)+
  theme(
    axis.text = element_text(size = 18, face = "bold"),
    axis.title = element_text(size = 20, face = "bold")
  )+
  xlab("Diet Source")

p12

tiff("EstBoxPlot_Fems.tiff", units="in", width=15, height=7, res=800)
p12
dev.off()


DietEst.Fems <- as.data.frame(bearsimmr_UnPrior_Fems$output$`1`$BUGSoutput$sims.list$p)

write.csv(DietEst.Fems, "DietEstFeamles.csv")

pal <- palette(viridis(5))

mytheme <- theme(
    axis.text = element_text(size = 18,face = "bold"),
    axis.title = element_text(size = 20, face = "bold"),
    panel.grid.major = element_line(color = "grey92"),
    panel.grid.minor = element_line(color = "grey96"),
    panel.background = element_rect(fill = "white"),
    axis.line = element_line(colour = "black",size = 1),
    axis.ticks = element_line(size = 1),
    )

p13 <- ggplot(data = DietEst.Fems)+
  geom_density(aes(Ants), color= pal[1], fill= pal[1], alpha=0.8)+
  geom_density(aes(Bilberry), color= pal[2], fill= pal[2], alpha=0.8)+
  geom_density(aes(Crowberry), color= pal[3], fill= pal[3], alpha=0.8)+
  geom_density(aes(Lingonberry), color= pal[4], fill= pal[4], alpha=0.8)+
  geom_density(aes(Moose), color= pal[5], fill= pal[5], alpha=0.8)+
  xlab("Diet Proportions")+
  ylab("Density")+
  theme(
    axis.text = element_text(size = 18,face = "bold"),
    axis.title = element_text(size = 20, face = "bold"),
    panel.grid.major = element_line(color = "grey92"),
    panel.grid.minor = element_line(color = "grey96"),
    panel.background = element_rect(fill = "white"),
    axis.line = element_line(colour = "black",size = 1),
    axis.ticks = element_line(size = 1),
    )+
  scale_y_continuous(expand = c(0,0), breaks = seq(0,200,25))+
  geom_text(label="Moose", x=0.048, y=105, size=8)+
  geom_text(label="Ants", x=0.07, y=70, size=8)+
  geom_text(label="Crowberry", x=0.152, y=15, size=8)+
  geom_text(label="Lingonberry", x=0.3, y=15, size=8)+
  geom_text(label="Bilberry", x=0.45, y=22, size=8)

p13

tiff("EstDenPlot_Fems.tiff", units="in", width=15, height=7, res=800)
p13
dev.off()



```


### Males

```{r}
bear.male <- subset(bear.si,sex!="F")
bear.male <- subset(bear.male, reprostatus.represent!="WithMother")

Mbearmix <- cbind(bear.male$C13.suess,bear.male$N15)
colnames(Mbearmix) = c('d13C','d15N')

bearsimmr_Male = simmr_load(
  mixtures=Mbearmix,
  source_names=bears_species_names,
  source_means=bears_species_means,
  source_sds=bears_species_sds,
  correction_means=bear_speciesTDF_means.Mikkelsen,
  correction_sds=bear_speciesc_sds,
  concentration_means = bear_species_conc
  )

p11 <- plot(bearsimmr_Male,
           xlab = expression(paste(delta^13, "C (\u2030)", sep = "")),
           ylab = expression(paste(delta^15, "N (\u2030)", sep = "")),
           mix_name = "Brown bears",
           title = ""
           )+
  scale_shape_manual(values = c(1,15,16,17,18,4))+
  theme(axis.text = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 23, face = "bold"),
        legend.text = element_text(size = 18),
        legend.title = element_blank())+
  scale_y_continuous(breaks = seq(-4,9,1),
                     labels = c("","","-2","","0","","2","",
                                "4","","6","","8",""))+
  coord_equal()

p11

bearsimmr_UnPrior_Males = simmr_mcmc(
  bearsimmr_Male,
  mcmc_control = list(
    iter=1000000, burn=500000, thin=500, n.chain=4
  )
)


summary(bearsimmr_UnPrior_Males, type = "diagnostics")

summary(bearsimmr_UnPrior_Males, type="statistics")
summary(bearsimmr_UnPrior_Males,'quantiles')
plot(bearsimmr_UnPrior_Males, type = "density")
plot(bearsimmr_UnPrior_Males, type = "matrix")


p14 <- plot(bearsimmr_UnPrior_Males,
  type = "boxplot",
  title = ""
)+
  theme(
    axis.text = element_text(size = 18, face = "bold"),
    axis.title = element_text(size = 20, face = "bold")
  )+
  xlab("Diet Source")

p14

tiff("EstBoxPlot_Males.tiff", units="in", width=15, height=7, res=800)
p14
dev.off()


DietEst.Males <- as.data.frame(bearsimmr_UnPrior_Males$output$`1`$BUGSoutput$sims.list$p)

write.csv(DietEst.Males,"DietEstMales.csv")

pal <- palette(viridis(5))

mytheme <- theme(
    axis.text = element_text(size = 18,face = "bold"),
    axis.title = element_text(size = 20, face = "bold"),
    panel.grid.major = element_line(color = "grey92"),
    panel.grid.minor = element_line(color = "grey96"),
    panel.background = element_rect(fill = "white"),
    axis.line = element_line(colour = "black",size = 1),
    axis.ticks = element_line(size = 1),
    )

p15 <- ggplot(data = DietEst.Males)+
  geom_density(aes(Ants), color= pal[1], fill= pal[1], alpha=0.8)+
  geom_density(aes(Bilberry), color= pal[2], fill= pal[2], alpha=0.8)+
  geom_density(aes(Crowberry), color= pal[3], fill= pal[3], alpha=0.8)+
  geom_density(aes(Lingonberry), color= pal[4], fill= pal[4], alpha=0.8)+
  geom_density(aes(Moose), color= pal[5], fill= pal[5], alpha=0.8)+
  xlab("Diet Proportions")+
  ylab("Density")+
  theme(
    axis.text = element_text(size = 18,face = "bold"),
    axis.title = element_text(size = 20, face = "bold"),
    panel.grid.major = element_line(color = "grey92"),
    panel.grid.minor = element_line(color = "grey96"),
    panel.background = element_rect(fill = "white"),
    axis.line = element_line(colour = "black",size = 1),
    axis.ticks = element_line(size = 1),
    )+
  scale_y_continuous(expand = c(0,0), breaks = seq(0,200,25))+
  geom_text(label="Moose", x=0.085, y=60, size=7)+
  geom_text(label="Ants", x=0.115, y=40, size=7)+
  geom_text(label="Crowberry", x=0.15, y=14, size=7)+
  geom_text(label="Lingonberry", x=0.27, y=12, size=7)+
  geom_text(label="Bilberry", x=0.61, y=21, size=7)

p15

tiff("EstDenPlot_Males.tiff", units="in", width=15, height=7, res=800)
p15
dev.off()
```

### Comparing Males & Females

```{r}
pal2 <- palette(viridis(30))

p16 <- ggplot()+
  geom_density(data = DietEst.Males,
               aes(Ants), color= pal2[1], fill= pal2[1], alpha=0.8)+
  geom_density(data = DietEst.Males,
               aes(Bilberry), color= pal2[5], fill= pal2[5], alpha=0.8)+
  geom_density(data = DietEst.Males,
               aes(Crowberry), color= pal2[9], fill= pal2[9], alpha=0.8)+
  geom_density(data = DietEst.Males,
               aes(Lingonberry), color= pal2[13], fill= pal2[13], alpha=0.8)+
  geom_density(data = DietEst.Males,
               aes(Moose), color= pal2[17], fill= pal2[17], alpha=0.8)+
   geom_density(data = DietEst.Fems,
               aes(Ants), color= pal2[3], fill= pal2[3], alpha=0.8)+
  geom_density(data = DietEst.Fems,
               aes(Bilberry), color= pal2[7], fill= pal2[7], alpha=0.8)+
  geom_density(data = DietEst.Fems,
               aes(Crowberry), color= pal2[11], fill= pal2[11], alpha=0.8)+
  geom_density(data = DietEst.Fems,
               aes(Lingonberry), color= pal2[15], fill= pal2[15], alpha=0.8)+
  geom_density(data = DietEst.Fems,
               aes(Moose), color= pal2[19], fill= pal2[19], alpha=0.8)+
  xlab("Diet Proportions")+
  ylab("Density")+
  theme(
    axis.text = element_text(size = 18,face = "bold"),
    axis.title = element_text(size = 20, face = "bold"),
    panel.grid.major = element_line(color = "grey92"),
    panel.grid.minor = element_line(color = "grey96"),
    panel.background = element_rect(fill = "white"),
    axis.line = element_line(colour = "black",size = 1),
    axis.ticks = element_line(size = 1),
    )+
  scale_y_continuous(expand = c(0,0), breaks = seq(0,200,25))+
  geom_text(label="Moose", x=0.085, y=60, size=7)+
  geom_text(label="Ants", x=0.115, y=40, size=7)+
  geom_text(label="Crowberry", x=0.15, y=14, size=7)+
  geom_text(label="Lingonberry", x=0.27, y=12, size=7)+
  geom_text(label="Bilberry", x=0.61, y=21, size=7)

p16

library(ggpattern)

p17 <- ggplot()+
  geom_density(data = DietEst.Males,
               aes(Ants), color= pal2[8], fill= pal2[8], alpha=0.7)+
  geom_density_pattern(data = DietEst.Males, aes(Ants),
                       pattern="stripe",
                       fill=pal2[8],
                       colour="black",
                       pattern_fill=pal2[8],
                       alpha=0.2,
                       pattern_density=0.7)+
  geom_density(data = DietEst.Fems,
               aes(Ants), color= pal2[1], fill= pal2[1], alpha=0.7, lwd=1.1)+
  xlab("Proportion Ants")+
  ylab("Density")+
  theme(
    axis.text = element_text(size = 16,face = "bold"),
    axis.title = element_text(size = 18, face = "bold"),
    panel.grid.major = element_line(color = "grey92"),
    panel.grid.minor = element_line(color = "grey96"),
    panel.background = element_rect(fill = "white"),
    axis.line = element_line(colour = "black",size = 1),
    axis.ticks = element_line(size = 1),
    )+
  scale_y_continuous(expand = c(0,0))+
  geom_text(label="Males", x=0.1, y=30, size=7)+
  geom_text(label="Females", x=0.06, y=77, size=7)

p17
  
tiff("EstDenPlot_AntsMF.tiff", units="in", width=15, height=7, res=800)
p17
dev.off()

p18 <- ggplot()+
  geom_density(data = DietEst.Males,
               aes(Bilberry), color= pal2[9], fill= pal2[9], alpha=0.7)+
  geom_density_pattern(data = DietEst.Males, aes(Bilberry),
                       pattern="stripe",
                       fill=pal2[9],
                       colour="black",
                       pattern_fill=pal2[9],
                       alpha=0.2,
                       pattern_density=0.7)+
  geom_density(data = DietEst.Fems,
               aes(Bilberry), color= pal2[2],
               fill= pal2[2], alpha=0.7, lwd=1.1)+
  xlab("Proportion Bilberry")+
  ylab("Density")+
  theme(
    axis.text = element_text(size = 16,face = "bold"),
    axis.title = element_text(size = 18, face = "bold"),
    panel.grid.major = element_line(color = "grey92"),
    panel.grid.minor = element_line(color = "grey96"),
    panel.background = element_rect(fill = "white"),
    axis.line = element_line(colour = "black",size = 1),
    axis.ticks = element_line(size = 1),
    axis.title.y = element_blank())+
  scale_y_continuous(expand = c(0,0))+
  geom_text(label="Males", x=0.1, y=30, size=7)+
  geom_text(label="Females", x=0.06, y=77, size=7)

p18
  
tiff("EstDenPlot_BilberryMF.tiff", units="in", width=15, height=7, res=800)
p18
dev.off()

p19 <- ggplot()+
  geom_density(data = DietEst.Males,
               aes(Crowberry), color= pal2[10], fill= pal2[10], alpha=0.7)+
  geom_density_pattern(data = DietEst.Males, aes(Crowberry),
                       pattern="stripe",
                       fill=pal2[10],
                       colour="black",
                       pattern_fill=pal2[10],
                       alpha=0.2,
                       pattern_density=0.7)+
  geom_density(data = DietEst.Fems,
               aes(Crowberry), color= pal2[3],
               fill= pal2[3], alpha=0.7, lwd=1.1)+
  xlab("Proportion Crowberry")+
  theme(
    axis.text = element_text(size = 16,face = "bold"),
    axis.title = element_text(size = 18, face = "bold"),
    panel.grid.major = element_line(color = "grey92"),
    panel.grid.minor = element_line(color = "grey96"),
    panel.background = element_rect(fill = "white"),
    axis.line = element_line(colour = "black",size = 1),
    axis.ticks = element_line(size = 1),
    axis.title.y = element_blank())+
  scale_y_continuous(expand = c(0,0))+
  geom_text(label="Males", x=0.1, y=30, size=7)+
  geom_text(label="Females", x=0.06, y=77, size=7)

p19
  
tiff("EstDenPlot_CrowberryMF.tiff", units="in", width=15, height=7, res=800)
p19
dev.off()

p20 <- ggplot()+
  geom_density(data = DietEst.Males,
               aes(Lingonberry), color= pal2[11], fill= pal2[11], alpha=0.7)+
  geom_density_pattern(data = DietEst.Males, aes(Lingonberry),
                       pattern="stripe",
                       fill=pal2[11],
                       colour="black",
                       pattern_fill=pal2[11],
                       alpha=0.2,
                       pattern_density=0.7)+
  geom_density(data = DietEst.Fems,
               aes(Lingonberry), color= pal2[4], fill= pal2[4],
               alpha=0.7, lwd=1.1)+
  xlab("Proportion Lingonberry")+
  ylab("Density")+
  theme(
    axis.text = element_text(size = 16,face = "bold"),
    axis.title = element_text(size = 18, face = "bold"),
    panel.grid.major = element_line(color = "grey92"),
    panel.grid.minor = element_line(color = "grey96"),
    panel.background = element_rect(fill = "white"),
    axis.line = element_line(colour = "black",size = 1),
    axis.ticks = element_line(size = 1))+
  scale_y_continuous(expand = c(0,0))+
  geom_text(label="Males", x=0.1, y=30, size=7)+
  geom_text(label="Females", x=0.06, y=77, size=7)

p20
  
tiff("EstDenPlot_LingonberryMF.tiff", units="in", width=15, height=7, res=800)
p20
dev.off()

p21 <- ggplot()+
  geom_density(data = DietEst.Males,
               aes(Moose), color= pal2[12], fill= pal2[12], alpha=0.7)+
  geom_density_pattern(data = DietEst.Males, aes(Moose),
                       pattern="stripe",
                       fill=pal2[12],
                       colour="black",
                       pattern_fill=pal2[12],
                       alpha=0.2,
                       pattern_density=0.7)+
  geom_density(data = DietEst.Fems,
               aes(Moose), color= pal2[5], fill= pal2[5], alpha=0.77, lwd=1.1)+
  xlab("Proportion Moose")+
  theme(
    axis.text = element_text(size = 16,face = "bold"),
    axis.title = element_text(size = 18, face = "bold"),
    panel.grid.major = element_line(color = "grey92"),
    panel.grid.minor = element_line(color = "grey96"),
    panel.background = element_rect(fill = "white"),
    axis.line = element_line(colour = "black",size = 1),
    axis.ticks = element_line(size = 1),
    axis.title.y = element_blank()
    )+
  scale_y_continuous(expand = c(0,0))
  
p21
  
tiff("EstDenPlot_MooseMF.tiff", units="in", width=15, height=7, res=800)
p21
dev.off()

library(gridExtra)

E <- grid.arrange(p17,p18,p19,p20,p21, ncol=3)
ggsave('EstDenPlot_GridMF.tiff',
       path = "C:\\Users\\amikk\\Documents\\Rprojects\\Diet-sensitivty-to-TDFs-and-priors",
       E,dpi=900, width= 300, height=190, units = "mm")


p22 <- plot(bearsimmr_TDF_Mikkelsen,
           xlab = "",
           ylab = expression(paste(delta^15, "N (\u2030)", sep = "")),
           mix_name = "Brown bears")+
  scale_shape_manual(values = c(1,15,16,17,18,4))+
  theme(axis.text = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 23, face = "bold"),
        legend.text = element_text(size = 18),
        plot.title = element_blank(),
        legend.position = "none")+
  scale_y_continuous(breaks = seq(-4,9,1),
                     labels = c("","","-2","","0","","2","",
                                "4","","6","","8",""))+
  coord_equal()+
  geom_text(label="A",size=8, y=8.5, x=-20.5)
p22


p23 <- plot(bearsimmr_Fems,
           xlab = expression(paste(delta^13, "C (\u2030)", sep = "")),
           ylab = expression(paste(delta^15, "N (\u2030)", sep = "")),
           mix_name = "Brown bears")+
  scale_shape_manual(values = c(1,15,16,17,18,4))+
  theme(axis.text = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 23, face = "bold"),
        legend.text = element_text(size = 18),
        legend.title = element_blank(),
        legend.position = "none",
        plot.title = element_blank(),
        axis.title.y = element_blank())+
  scale_y_continuous(breaks = seq(-4,9,1),
                     labels = c("","","-2","","0","","2","",
                                "4","","6","","8",""))+
  coord_equal()+
  geom_text(label="B",size=8, y=8.5, x=-20.5)

p23

p24 <- plot(bearsimmr_Male,
           xlab = "",
           ylab = expression(paste(delta^15, "N (\u2030)", sep = "")),
           mix_name = "Brown bears")+
  scale_shape_manual(values = c(1,15,16,17,18,4))+
  theme(axis.text = element_text(size = 18, face = "bold"),
        axis.title = element_text(size = 23, face = "bold"),
        legend.text = element_text(size = 18),
        legend.title = element_blank(),
        plot.title = element_blank(),
        axis.title.y = element_blank(),
        legend.position = "none")+
  scale_y_continuous(breaks = seq(-4,9,1),
                     labels = c("","","-2","","0","","2","",
                                "4","","6","","8",""))+
  coord_equal()+
  geom_text(label="C",size=8, y=8.5, x=-20.5)

p24


F <- grid.arrange(p22,p23,p24, ncol=3)
ggsave('MixingSpace_3.tiff',
       path = "C:\\Users\\amikk\\Documents\\Rprojects\\Diet-sensitivty-to-TDFs-and-priors",
       F,dpi=900, width= 300, height=190, units = "mm")

```

